Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-14 talk/pepm2011/paper.tex
A short paragraph about Psyco in Related Work.
expand for clarification

One problem to the successful application of the allocation removal techniques
described in the previous sections is the presence of frame-introspection
features in many dynamic languages. Languages such as Python and Smalltalk allow
the programmer to get access to the frames object that the interpreter uses to
store local variables. This is a useful feature, as makes the implementation of
a debugger possible in Python without needing much support from the VM level. On
the other hand, it severely hinders the effectiveness of allocation removal,
because every time an object is stored into a local variable, it is stored into
the frame-object, which makes it escape.

XXX [fijal] Well, in fact that's *a lot* of support for debugger from the VM.
The support is just generic instead of specific.

This problem is solved by making it possible to the interpreter author to add
some hints into the source code to declare instances of one class as frame
objects. The JIT will then fill these objects only lazily when they are actually
accessed (\eg because a debugger is used). Therefore in the common case, nothing
is stored into the frame objects, making the problem of too much escaping go
away. This is a common approach in VM implementations \cite{miranda_context_1999}, the only
novelty in our approach lays in its generality, because most other JITs are just
specifically written for one particular language.

@@

@@

XXX Psyco \cite{rigo_representation-based_2004}

>>>>

One problem to the successful application of the allocation removal
techniques described in the previous sections is the presence of
frame-introspection features in many dynamic languages. Languages such
as Python and Smalltalk allow the programmer to get access to the
frames object that the interpreter uses to store local variables. This
is a useful feature, as makes the implementation of a debugger
possible in Python without needing much explicit support from the VM
level. On the other hand, it severely hinders the effectiveness of
allocation removal, because every time an object is stored into a
local variable, it is stored into the frame-object, which makes it
escape.

This problem is solved by making it possible to the interpreter author
to add some hints into the source code to declare instances of one
class as frame objects needing special treatment. The JIT will then
fill these objects only lazily when they are actually accessed (\eg
because a debugger is used) using values from JIT-emitted code runtime
locations (typically the CPU stack). Therefore in the common case,
nothing is stored into the frame objects, making the problem of too
much escaping go away. This is a common approach in VM implementations
\cite{miranda_context_1999}, the only novelty in our approach lays in
its generality, all the delicate support code for this is generated, because
most other JITs are just specifically written for one particular
language.

@@

Psyco \cite{rigo_representation-based_2004} is a (non-tracing) JIT for Python
that fully implements allocation removal (the \emph{static} objects are called
\emph{virtual} objects).  It is a hand-written extension module for CPython.
Historically, PyPy's JIT can be seen as some successor of Psyco for a
general context.

@@
